---
title: "msPVA : An R implementation of count-based multi-site population viability analysis"
output: 
  html_document:
    highlight: haddock
    keep_md: yes
    theme: readable
---
[![DOI](https://zenodo.org/badge/19481/cmartin/msPVA.svg)](https://zenodo.org/badge/latestdoi/19481/cmartin/msPVA)
[![Build Status](https://travis-ci.org/cmartin/msPVA.svg?branch=master)](https://travis-ci.org/cmartin/msPVA)
[![Coverage Status](https://coveralls.io/repos/cmartin/msPVA/badge.svg?branch=master&service=github)](https://coveralls.io/github/cmartin/msPVA?branch=master)

This package implements a stochastic simulation for a count-based multi-site population viability analysis as described in chapter 11 of Quantitative Conservation Ecology (Morris & Doak, 2002).

Code is highly inspired from the MatLab implementation described in the book.

## Installation
You have two options here.

### Install package from github.com
This allows you to easily install updates, have access to the function help files etc.
```{r eval=FALSE}
library(devtools)
devtools::install_github("cmartin/msPVA")
library(msPVA)
```

### Or you download the source file and run it directly
In this case, you need download [the raw R file](https://raw.githubusercontent.com/cmartin/msPVA/master/R/MultisitePVA.R)
and then execute : 
```{r eval=FALSE}
source("MultisitePVA.R")
```
This is a quicker, but dirtier way!

## Try some examples

With the same Clapper rail data as in Morris & Doak

### Definining parameters manually : 
```{r echo=FALSE}
library(msPVA)
library(knitr)
read_chunk("inst/examples/ms.Example.R")
read_chunk("inst/examples/ss.Example.R")
read_chunk("inst/examples/params.Example.R")
```
```{r msExample, comment=NA}
```

### Or ask the package to calculate most of them automatically from a time-series : 
With a two-populations polar bear time series from [Stirling et al. 2004](http://arctic.journalhosting.ucalgary.ca/arctic/index.php/arctic/article/view/479/509)
```{r paramsExample, comment=NA}
```

## Run it for a single population
```{r ssExample, comment=NA}
```

## Read the book
If you are to use this package, I highly recommend that you first read chapter 11 from [Quantitative Conservation Ecology (Morris & Doak, 2002)](http://www.sinauer.com/quantitative-conservation-biology-theory-and-practice-of-population-viability-analysis.html), so you understand limitations and assumptions from the underlying model. Managing animal populations should not be taken lightly.

## Problems
Please report any bugs to the [GitHub issue tracker](https://github.com/cmartin/msPVA/issues) and write any questions to <charles.martin1@uqtr.ca>

## Citation
If this code is useful to you, please cite as : 

```{r echo=FALSE, comment=NA}
cat(citation("msPVA")$textVersion)
```